Supernovae Time Profiles as a Probe of New Physics at Neutrino Telescopes

### 2024-03-13 paper/main.tex
Updates from Overleaf

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\textbf{\textit{Introduction}}---
The Standard Model (SM) is a remarkable but incomplete theory. 
Puzzles such as the non-vanishing neutrino mass, the origin of observed matter-antimatter asymmetry, and the nature of dark matter, among others, seek explanations beyond the SM (BSM).
There are also no firmly established clues on the energy scale at which the new particles should appear.
To this end, numerous BSM searches across a wide range of scales have been performed, from collider searches at high energies~\cite{Nath:2010zj} to studies of cosmic microwave background at temperatures near absolute zero~\cite{Baumann:2015rya}.
Supernovae release most of their energy in neutrinos, offering unique opportunities to test BSM physics via neutrino interactions.
The duration of the burst~\cite{Kamiokande-II:1987idp,Bionta:1987qt,Baksan} and inferred total energy of neutrinos~\cite{Loredo:2001rx,Pagliaroli:2008ur,Huedepohl2010} from SN 1987A---the only supernova observed in neutrinos---have already given tentative constraints on the energies that could be taken away by any light BSM particle copiously produced in the interior of a star.
Indeed, many different BSM scenarios models have been constrained in this way, see \textit{e.g.} Refs.~\cite{Raffelt:2011nc,Arguelles:2016uwb,Suliga:2020vpz,Lucente:2021hbp,Caputo:2022rca,Caputo:2021rux,PhysRevD.100.083002,DeRocco:2019njg,Kazanas:2014mca,Magill:2018jla}.

If, additionally, BSM particles emitted from supernovae can decay to neutrinos en route to Earth, even stronger limits can be set, as shown, for instance, in Ref.~\cite{Fiorillo:2022cdq} for Majoron-like bosons and in Ref.~\cite{Brdar:2023tmi} for a realization featuring a neutrino magnetic moment portal.
Specifically, unlike neutrinos inside the supernova core, produced BSM states can stream out without further interactions.
Consequently, they exit with higher energies than neutrinos, typically around $\mathcal{O}(100)$~MeV.
Provided such particles decay to neutrinos, strong constraints can be set from the fact that neutrinos of $\mathcal{O}(100)$~MeV were not recorded from SN 1987A~\cite{Fiorillo:2022cdq, Brdar:2023tmi}.
Such strategies were utilized for neutrino experiments operating during the SN 1987A event. Sensitivity projections for  DUNE~\cite{DUNE:2015lol} and Hyper-Kamiokande~\cite{Hyper-Kamiokande:2018ofw} in light of anticipated galactic supernova event were also performed.

\begin{figure}[t!]
    \centering
    \includegraphics[width=0.47\textwidth]{figures/Supernova_Diagram_Draft-1.pdf}
    \caption{\textbf{The decay geometry for Majorons produced in supernova.}}
    \label{fig:geometry}
\end{figure}

The IceCube Neutrino Observatory has pioneered the detection of neutrinos at $\mathcal{O}$(PeV) energies~\cite{IceCube:2014stg,IceCube:2018cha} and has identified several specific astrophysical neutrino sources~\cite{IceCube:2018cha,IceCube:2023ame,IceCube:2022der}.
IceCube is also expected to be able to detect the next galactic supernova~\cite{Kopke_2011}.
In fact, a search for MeV neutrinos from optically obscured galactic supernovae was recently performed in \cite{IceCube:2023ogt} as well as the search for temporal correlation of MeV neutrino events with fast radio bursts \cite{IceCube:2019acm}.
In both cases, the observable is a collective rise in all photomultiplier rates on top of the background noise  within a certain time window, and it turns out that IceCube is sensitive to intervals as short as $\mathcal{O}(0.01)$ seconds. 
In this letter, we demonstrate that such precise time resolution will play a crucial role in testing new physics from supernovae.

Once produced in supernovae, the BSM particles can be nonrelativistic and travel slowly, causing time delays for secondary neutrinos. However, it is also possible that new states are rather light, implying that they may even exit the supernova at earlier times with respect to certain flavors of SM neutrinos.
Such an early BSM signal, which to the best of our knowledge has not been previously considered, can, for instance, be induced by light BSM states produced from $\nu_e$ scattering inside the supernova; they would then subsequently decay to $\bar{\nu}_e$, which chiefly induces the signal at IceCube. 
This would happen at early stages of a supernova, around the neutralization burst, when $\nu_e$ are present and antineutrinos have not been produced yet.
Such novel timing patterns generically exist in models with new weakly interacting particles, such as the aforementioned Majoron model~\cite{Fiorillo:2022cdq} and the neutrino magnetic moment portal~\cite{Magill:2018jla,Brdar:2020quo}.
Via the timing measurement, we will show that neutrino telescopes will be able to provide powerful constraints on the BSM parameter space in association with forthcoming galactic supernova events.\\

@@

\textbf{\textit{The Model}}---
Let us first consider the Majoron model, studied in Ref.~\cite{Fiorillo:2022cdq}, where supernova 1987A constraints were derived.
The relevant part of the Lagrangian reads
\begin{align}
\mathcal{L} \supset -g_{\alpha\beta} \nu_\alpha \nu_\beta \phi + \text{h.c.} - m_\phi \phi\phi^*\,,
\end{align}
where $\phi$ is the (pseudo)scalar, $m_\phi$ is the respective mass, and $g$ parametrizes interaction strength between $\phi$ and neutrinos. We assume flavor universal interaction and hence $g_{\alpha\beta}\equiv g_\phi \, \delta_{\alpha\beta}$. 
Majorons are produced from (anti)neutrino coalescence in the star and subsequently decay to a pair of (anti)neutrinos resulting in BSM (anti)neutrino flux emitted from the star.
The total decay width is given by $\Gamma_\phi = 3g^2 m_\phi/16 \pi$.

\begin{figure}[t!]
    \centering
    \includegraphics[width=0.47\textwidth]{figures/majoran_fluxes.pdf}
    \caption{\textit{\textbf{Flux of $\bar{\nu}_{e}$ from SM production and two Majoron hypotheses}}.
    We choose the parameters $g_\phi = 10^{-10.2}$ and $g_\phi = 10^{-11.8}$ for $m_\phi = \unit[1]{MeV}$ and $m_\phi = \unit[200]{MeV}$, respectively.
    }
    \label{fig:fluxes}
\end{figure}

We will set supernova bounce time $t_{\rm bounce}$ as $t = 0$.
Using data from the simulation of $8.8 M_\odot$ progenitor star \cite{Huedepohl2010} and including effects of neutrino oscillations assuming normal mass ordering \cite{Dighe:1999bi}, we calculated standard neutrino fluxes at Earth. Following Ref.~\cite{Fiorillo:2022cdq}, we also calculated the flux of emitted Majorons.
At $t \leq \unit[0.05]{s}$, Majorons are mainly produced through $\nu_e$ and $\nu_x$ coalescence where $\nu_x$ includes (anti)neutrinos of $\tau$ and $\mu$ flavor. After $\unit[0.05]{s}$, the flux of Majorons decreases with the total flux of neutrinos of all flavors. Thus we found that the Majoron flux peaks around $t \sim \unit[0.05]{s}$. We further cross checked the agreement with Ref.~\cite{Fiorillo:2022cdq} by reproducing their bound from supernova energy loss.

Very weakly coupled Majorons immediately stream out after the production.
On their way to Earth, Majorons of energy $E_\phi$ will travel at the speed of $\beta$ for a distance of $L_1$, then decay to (anti)neutrinos of energy $E_\nu$ at an emission angle $\cos\alpha = (2 E_\phi E_{\nu} - m^2_\phi)/(2E_{\nu}E_\phi\beta)$.
Daughter (anti)neutrinos will travel for a distance $L_2$ and reach Earth at the angle $\theta$ satisfying $D_{\rm SN} \sin\theta = L_1 \sin\alpha$ for a supernova that is $D_{\rm SN}$ away (see the decay geometry in \cref{fig:geometry}).
The time delay, $\delta t$, relative to a relativistic particle moving along the line-of-sight path from a supernova to reach the Earth is given by
\cite{Jaeckel:2017tud}
\begin{align}
    \delta t = \left(\frac{L_1}{\beta}-L_1\right) + (L_1 + L_2 -D_{\rm SN})\,,
    \label{eq:deltatexa}
\end{align}
where we have split the contribution to $\delta t$ into two parts: the first part comes from the slow-moving Majorons before their decay, and the second part is due to the detour from the line-of-sight path at the given emission angle $\alpha$.

We will consider a supernova event that happens in the galaxy at a distance $D_{\rm SN}=\unit[10]{kpc}$, which is not unlikely~\cite{Reed:2005en,Rozwadowska:2020nab}. To get the characteristic value of $\delta t$, we can take $L_1 = L_\phi$ with $L_\phi = (E_\phi/m_\phi) \Gamma^{-1}_\phi \beta$ being the decay length of $\phi$.
As for the parameter regions we considered, $(D_{\rm SN} -D_{\rm SN} \cos\theta) \ll \unit[10^{-3}]{s}$, $\delta t$ can be approximated as
\begin{align}
    \delta t &\simeq L_\phi\left(\frac{1}{\beta} -1\right) + L_\phi(1-\cos\alpha) = \frac{8\pi}{3 E_{\nu}g_\phi^2}\,,
    \label{eq:deltat}
\end{align}
which is roughly independent of Majoron energy $E_\phi$.
%However, at small values of $E_\phi$, the main contribution to $\delta t$ comes from the first term due to the slow-moving Majorons, while at larger values of $E_\phi$, from the second term due to a larger detour given the increase of $L_\phi$ over the decrease of $(1-\cos\alpha)$.

\begin{figure*}[t]
    \centering
    \includegraphics[width=0.95\textwidth]{figures/hits_and_likelihood.pdf}
    \caption{\textbf{\textit{Timing profile of hits and test statistic as a function of the timing window.}}
    The hit rate from the SM neutrino flux peaks at around 0.1~s, coinciding with the peak emission of $\bar{\nu}_{e}$ from the SN.
    The shaded regions correspond to the timing windows that give peak sensitivity to the particular model. These are the same benchmarks in the parameter space as in Fig.~\ref{fig:fluxes}.
    In the right panel, we show the test statistic as a function of the starting time and length of the window for the heavy Majoron benchmark ($m_\phi=200$ MeV).
    The relatively broad distribution of high test-statistic values shows that this analysis technique does not require extremely precise reconstruction of the $t_{\mathrm{bounce}}$.
    }
    \label{fig:hits_and_likelihood}
\end{figure*}

We can obtain the flux of daughter $\nu$ by considering decays that occur only at $L_1\geq R^{\nu}_{\rm SN}$.
As the smallest decay distance beyond which the daughter neutrino can escape the explosion unperturbed is roughly the radius of the core, we take $R^{\nu}_{\rm SN} = \unit[30]{km}$.
We will also limit our analysis to a maximal time window of $\unit[100]{s}$ after supernova bounce.
This time window is larger than typical values of time delay arising in the model.

@@

\textbf{\textit{Detector Response and Statistical Treatment}}---The IceCube Neutrino Observatory comprises 5,160 light-detecting digital optical modules (DOMs) buried in a cubic kilometer of the deep, transparent Antarctic ice sheet.
The DOMs are arranged on 86 strings of 60 DOMs each, with an interstring distance of 125~m (70~m for the eight centermost strings).
These DOMs detect the photons emitted by the charged byproducts of a neutrino interaction in or near the instrumented volumen.
This enables IceCube to resolve individual neutrinos with energies ranging between a few GeV and a few PeV.
The neutrinos produced by supernovae are far below this threshold and cannot be individually resolved; however, the immense number of neutrinos produced in a supernova increases the single-photon rate of the detector.
This dramatic increase in the rate of photons can be distinguished from the background caused by dark noise and radioactive activity in the DOM glass to enable the detection of galactic supernovae events~\cite{Griswold:2023iwz}.

New physics will affect the development of a supernova and distort the temporal structure of the photon signal seen in the IceCube detector.
Thus, one may look for excess events in certain time intervals as evidence of this new physics.
In this work, we simulate the light curve produced by a standard supernova with a progenitor mass of $8.8~M_{\odot}$ and those produced by different BSM scenarios mentioned previously; see the left-hand panel of Fig.~\ref{fig:hits_and_likelihood} for light curves of the same two benchmarks from \cref{fig:fluxes}.

Since more than 93\% of the photons detected in IceCube originate in IBD interactions, the SM-only event rate peaks at around 0.1~s.
BSM scenarios, on the other hand, can give rise to an early or delayed signal depending on the particulars of the model.
In essence, Majorons can be generated before the peak of the SM $\bar{\nu}_{e}$ flux, and as light Majorons can travel nearly relativistically, they can escape the supernovae with negligible time delay and then decay to an all-flavor flux of neutrinos, producing early $\bar{\nu}_{e}$ fluxes before the SM one.
The higher-mass Majorons, on the other hand, acquire a time delay from traveling sub-luminally and at a large detour, thus producing a later signal with sizable time delay.
We use the \texttt{ASTERIA}~\cite{spencer_griswold_2020_3926835} package to simulate the detector response to the different supernova scenarios and the thermal and radioactive noise in the DOMs.
This package simulates light yields from coherent $\nu_{e}$ scattering off electrons in the ice, IBD of $\bar{\nu}_{e}$ on nuclei in the ice, and charged- and neutral-current interactions for $\nu_{\alpha}$. In addition to photons produced in neutrino interactions, it also simulates photons from thermal noise in the IceCube PMTs and from radioactive decay in the pressure glass housing.
In order to interface with \texttt{ASTERIA}, we use the \texttt{SNEWPY}~\cite{SNEWS:2021ewj} package and, in particular, the \texttt{ParametrizedFlux} object to parametrize the flux with the ``pinching'' factor~\cite{Keil:2002in,SNEWS:2021ewj}.

@@

\textbf{\textit{Results and Discussions}}---
\begin{figure}[t!]
    \centering
    \includegraphics[width=0.47\textwidth]{figures/majoran_sensitivity}
    \caption{\textbf{\textit{Exclusion sensitivities for the Majoron case.}}
    The lines on this plot show the exclusion sensitivity for IceCube, IceCube Gen-2, Hyper-Kamiokande and DUNE.
    %The sensitivities shown as solid lines are computed from this work, whereas the dashed lines are computed following~\cite{Brdar:2023tmi}.
    Additionally, we also show the reaches by supernova cooling and by the lack of observation of high-energy neutrinos from SN1987A as black lines. %\cite{Fiorillo:2022cdq}, with the width of the shading region quantifying uncertainties from supernova modeling.
    }
    \label{fig:sensitivity}
\end{figure} 
Following our strategies above, we obtain the expected exclusion limits for various BSM scenarios from no observed excess of hits on top of background noise and standard neutrino flux.
The expected exclusion limits are presented in \cref{fig:sensitivity} for the Majoron case considering both IceCube and IceCube-Gen2. To illustrate the broad physics application of the analysis, we also show results for the sterile neutrino case with magnetic portal in the Supplemental Materials.
Current limits \cite{Fiorillo:2022cdq} from the energy loss requirement, and non-observation of high-energy supernova neutrinos from SN 1987A are also shown in \cref{fig:sensitivity} for the Majoron case. With future supernova explosions, IceCube can improve current limits by one order of magnitude. 
Moreover, while current limits from time-integrated observables follow the behavior of $g_\phi\propto m^{-1}_\phi$ as Majoron production rate is proportional to $(g_\phi m_\phi)^2$, IceCube can provide a stronger limit than this by focusing on the time window, especially before the peak of SM $\bar{\nu}_e$ fluxes, to reduce SM condemnations.

@@

There are existing uncertainties from different modeling of supernova explosions (see e.g. Ref.~\cite{li2023old}) as well as discrepancies between SN 1987A neutrino data and modeling. Nevertheless, we point out that even the uncertainties of around $2\sigma$  will only change our exclusion limits by a small factor.
We make the code used in this work publicly available so that the impact of supernova modeling uncertainties on the sensitivity of neutrino telescopes to BSM scenarios can be further explored.

\textbf{\textit{Summary}}---
In this letter, we have demonstrated that neutrino telescopes such as IceCube can provide unprecedented probes of new physics, such as Majoron-like bosons and neutrino magnetic moment portal, by measuring supernova time profiles to a window as short as $\mathcal{O}(0.01)$ s. Specifically, signatures outside the SM neutrino time window, including both early and delayed signals, can be effectively probed by neutrino telescopes. This analysis can be extended to test other BSM realizations including, but not limited to, sterile neutrinos with a mixing portal. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\textit{Acknowledgements}---
We would like to thank Thomas Janka and Daniel Kresse for providing the data from Garching supernovae simulations (available at \url{https://wwwmpa.mpa-garching.mpg.de/supernovaarchive/}) in machine-readable form, and also for stimulating discussions. Y.-Y. L is supported by the NSF of China through Grant No. 12305107, 12247103.
CAA are supported by the Faculty of Arts and Sciences of Harvard University, the National Science Foundation (NSF), the NSF AI Institute for Artificial Intelligence and Fundamental Interactions (IAIFI), the Research Corporation for Science Advancement, 
and the David \& Lucile Packard Foundation.
CAA and JL were partially supported by the Alfred P. Sloan Foundation for part of this work.
JL acknowledges support from the Belgian American Educational Foundation.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\textbf{\textit{Introduction}}---
The Standard Model (SM) is a remarkable but incomplete theory. 
Puzzles such as the non-vanishing neutrino mass, the origin of observed matter-antimatter asymmetry, and the nature of dark matter, among others, seek explanations beyond the SM (BSM).
There are also no firmly established clues on the energy scale at which the new particles should appear.
To this end, numerous BSM searches across a wide range of scales have been performed, from collider searches at high energies~\cite{Nath:2010zj} to studies of cosmic microwave background at temperatures near absolute zero~\cite{Baumann:2015rya}.

Supernovae (SNe) release most of their energy in neutrinos, offering unique opportunities to test BSM physics via neutrino interactions.
The duration of the burst~\cite{Kamiokande-II:1987idp,Bionta:1987qt,Baksan} and inferred total energy of neutrinos~\cite{Loredo:2001rx,Pagliaroli:2008ur,Huedepohl2010} from SN 1987A---the only SN observed in neutrinos---have already given tentative constraints on the energies that could be taken away by any light BSM particle copiously produced in the interior of a star.
Indeed, many different BSM scenarios have been constrained in this way, see \textit{e.g.} Refs.~\cite{Raffelt:2011nc,Arguelles:2016uwb,Suliga:2020vpz,Lucente:2021hbp,Caputo:2022rca,Caputo:2021rux,PhysRevD.100.083002,DeRocco:2019njg,Kazanas:2014mca,Magill:2018jla}. If, additionally, BSM particles emitted from SNe can decay to neutrinos en route to Earth, even stronger limits can be set, as shown, for instance, in Ref.~\cite{Fiorillo:2022cdq} for Majoron-like bosons and in Ref.~\cite{Brdar:2023tmi} for a realization featuring a neutrino magnetic moment portal.
Specifically, unlike neutrinos inside the SN core, produced BSM states can stream out without further interactions.
Consequently, they exit with higher energies than neutrinos, typically around $\mathcal{O}(100)$~MeV.
Provided such particles decay to neutrinos, strong constraints can be set from the fact that neutrinos of $\mathcal{O}(100)$~MeV were not recorded from SN 1987A~\cite{Fiorillo:2022cdq, Brdar:2023tmi}.
Such strategies were utilized for neutrino experiments operating during the SN 1987A event. Sensitivity projections for  DUNE~\cite{DUNE:2015lol} and Hyper-Kamiokande~\cite{Hyper-Kamiokande:2018ofw} in light of anticipated galactic SN event were also performed.

The IceCube Neutrino Observatory has pioneered the detection of neutrinos at $\mathcal{O}$(PeV) energies~\cite{IceCube:2014stg,IceCube:2018cha} and has identified several specific astrophysical neutrino sources~\cite{IceCube:2018cha,IceCube:2023ame,IceCube:2022der}.
IceCube is also expected to be able to detect the next galactic SN~\cite{Kopke_2011}.
In fact, a search for MeV neutrinos from optically obscured galactic SNe was recently performed in \cite{IceCube:2023ogt} as well as the search for temporal correlation of MeV neutrino events with fast radio bursts \cite{IceCube:2019acm}.
In both cases, the observable is a collective rise in all photomultiplier rates on top of the background noise  within a certain time window, and it turns out that IceCube is sensitive to intervals as short as $\mathcal{O}(0.01)$ seconds. 
In this Letter, we will demonstrate that such precise time resolution will play a crucial role in testing new physics from SNe.

Once produced in SNe, the BSM particles can be nonrelativistic and travel slowly, causing time delays for secondary neutrinos. However, it is also possible that new states are rather light, implying that they may even exit the SN at earlier times with respect to certain flavors of SM neutrinos.
Such an early BSM signal, which to the best of our knowledge has not been previously considered, can, for instance, be induced by light BSM states produced from $\nu_e$ scattering inside the SN; they would then subsequently decay to $\bar{\nu}_e$, which chiefly induces the signal at IceCube. 
This would happen at early stages of a SN, around the neutralization burst, when $\nu_e$ are present and antineutrinos have not been produced yet.
Such novel timing patterns generically exist in models with new weakly interacting particles, such as the aforementioned Majoron model~\cite{Fiorillo:2022cdq} and the neutrino magnetic moment portal~\cite{Magill:2018jla,Brdar:2020quo}.

In this work we will show that, by using the timing measurements, neutrino telescopes will be able to provide powerful constraints on the BSM parameter space in association with forthcoming galactic SN events.\\
%Via the timing measurement, we will show that neutrino telescopes will be able to provide powerful constraints on the BSM parameter space in association with forthcoming galactic SN events.\\

@@

\textbf{\textit{The Model}}---
Let us first consider the Majoron model, studied in Ref.~\cite{Fiorillo:2022cdq}, where SN 1987A constraints were derived.
The relevant part of the Lagrangian reads
\begin{align}
\mathcal{L} \supset -g_{\alpha\beta} \nu_\alpha \nu_\beta \phi + \text{h.c.} - m_\phi \phi\phi^*\,,
\end{align}
where $\phi$ is the (pseudo)scalar, $m_\phi$ is the respective mass, and $g$ parametrizes interaction strength between $\phi$ and neutrinos. We assume flavor universal interaction and hence $g_{\alpha\beta}\equiv g_\phi \, \delta_{\alpha\beta}$. 
Majorons are produced from (anti)neutrino coalescence in the star and subsequently decay to a pair of (anti)neutrinos resulting in BSM (anti)neutrino flux emitted from the star.
The total decay width is given by $\Gamma_\phi = 3g^2 m_\phi/16 \pi$.

We will set SN bounce time $t_{\rm bounce}$ as $t = 0$.
Using data from the simulation of $8.8 M_\odot$ progenitor star \cite{Huedepohl2010} and including effects of neutrino oscillations assuming normal mass ordering \cite{Dighe:1999bi}, we calculated standard neutrino fluxes at Earth. Following Ref.~\cite{Fiorillo:2022cdq}, we also calculated the flux of emitted Majorons.
At $t \leq \unit[0.05]{s}$, Majorons are mainly produced through $\nu_e$ and $\nu_x$ coalescence where $\nu_x$ includes (anti)neutrinos of $\tau$ and $\mu$ flavor. After $\unit[0.05]{s}$, the flux of Majorons decreases with the total flux of neutrinos of all flavors. Thus we found that the Majoron flux peaks around $t \sim \unit[0.05]{s}$. We further cross checked the agreement with Ref.~\cite{Fiorillo:2022cdq} by reproducing their bound from SN energy loss.

\begin{figure}[t!]
\vspace{0.27cm}
    \centering
    \includegraphics[width=0.48\textwidth]{figures/Supernova_Diagram_Draft-2}
    \caption{\textbf{The decay geometry for Majorons produced in supernova.}}
    \label{fig:geometry}
\end{figure}

Very weakly coupled Majorons immediately stream out after being produced inside the star.
On their way to Earth, Majorons of energy $E_\phi$ will travel at the speed of $\beta$ for a distance of $L_1$, then decay to (anti)neutrinos of energy $E_\nu$ at an emission angle $\cos\alpha = (2 E_\phi E_{\nu} - m^2_\phi)/(2E_{\nu}E_\phi\beta)$. Daughter (anti)neutrinos will travel for a distance $L_2$ and reach Earth at the angle $\theta$ satisfying $D_{\rm SN} \sin\theta = L_1 \sin\alpha$ for a SN that is $D_{\rm SN}$ away (see the decay geometry in \cref{fig:geometry}).
The time delay, $\delta t$, relative to a relativistic particle moving along the line-of-sight path from a SN to reach the Earth is given by
\cite{Jaeckel:2017tud}
\begin{align}
    \delta t = \left(\frac{L_1}{\beta}-L_1\right) + (L_1 + L_2 -D_{\rm SN})\,,
    \label{eq:deltatexa}
\end{align}
where we have split the contribution to $\delta t$ into two parts: the first part comes from the slow-moving Majorons before their decay, and the second part is due to the detour from the line-of-sight path at the given emission angle $\alpha$.

\begin{figure}[t!]
    \centering
    \includegraphics[width=0.47\textwidth]{figures/majoran_fluxes.pdf}
    \caption{\textit{\textbf{Flux of $\bar{\nu}_{e}$ from SM production and two Majoron hypotheses at Earth}}.
    We choose the parameters $g_\phi = 10^{-10.2}$ and $g_\phi = 10^{-11.8}$ for $m_\phi = \unit[1]{MeV}$ and $m_\phi = \unit[200]{MeV}$, respectively.
    }
    \label{fig:fluxes}
\end{figure}

We will consider a SN event that happens in the galaxy at a distance $D_{\rm SN}=\unit[10]{kpc}$, which is not unlikely~\cite{Reed:2005en,Rozwadowska:2020nab}. To get the characteristic value of $\delta t$, we can take $L_1 = L_\phi$ with $L_\phi = (E_\phi/m_\phi) \Gamma^{-1}_\phi \beta$ being the decay length of $\phi$.
As for the parameter regions we considered, $(D_{\rm SN} -D_{\rm SN} \cos\theta) \ll \unit[10^{-3}]{s}$, $\delta t$ can be approximated as
\begin{align}
    \delta t &\simeq L_\phi\left(\frac{1}{\beta} -1\right) + L_\phi(1-\cos\alpha) = \frac{8\pi}{3 E_{\nu}g_\phi^2}\,,
    \label{eq:deltat}
\end{align}
which is roughly independent of Majoron energy $E_\phi$.

\begin{figure*}[t]
    \centering
    \includegraphics[width=0.95\textwidth]{figures/hits_and_likelihood.pdf}
    \caption{\textbf{\textit{Timing profile of hits and test statistic as a function of the timing window.}}
    In the left panel, we show the hit rate from the SM neutrino flux which peaks at around 0.1~s, coinciding with the peak emission of $\bar{\nu}_{e}$ from the SN. The shaded regions correspond to the timing windows that give peak sensitivity to the particular model. These are the same benchmarks in the parameter space as in Fig.~\ref{fig:fluxes}.
    In the right panel, we show the test statistic as a function of the starting time and length of the window for the heavy Majoron benchmark ($m_\phi=200$ MeV).
    The relatively broad distribution of high test-statistic values shows that this analysis technique does not require extremely precise reconstruction of the $t_{\mathrm{bounce}}$.
    }
    \label{fig:hits_and_likelihood}
\end{figure*}

We can obtain the flux of daughter $\nu$ by considering decays that occur only at $L_1\geq R^{\nu}_{\rm SN}$.
As the smallest decay distance beyond which the daughter neutrino can escape the explosion unperturbed is roughly the radius of the core, we take $R^{\nu}_{\rm SN} = \unit[30]{km}$.
We will also limit our analysis to a maximal time window of $\unit[100]{s}$ after SN bounce.
This time window is larger than typical values of time delay arising in the model.

@@

\textbf{\textit{Detector Response and Statistical Treatment}}---The IceCube Neutrino Observatory comprises 5,160 light-detecting digital optical modules (DOMs) buried in a cubic kilometer of the deep, transparent Antarctic ice sheet.
The DOMs are arranged on 86 strings of 60 DOMs each, with an interstring distance of 125~m (70~m for the eight centermost strings).
These DOMs detect the photons emitted by the charged byproducts of a neutrino interaction in or near the instrumented volume.
This enables IceCube to resolve individual neutrinos with energies ranging between a few GeV and a few PeV.
The neutrinos produced by SNe are far below this threshold and cannot be individually resolved; however, the immense number of neutrinos produced in a SN increases the single-photon rate of the detector.
This dramatic increase in the rate of photons can be distinguished from the background caused by dark noise and radioactive activity in the DOM glass to enable the detection of galactic SN events~\cite{Griswold:2023iwz}.

New physics will affect the development of a SN and distort the temporal structure of the photon signal seen in the IceCube detector.
Thus, one may look for excess events in certain time intervals as evidence of this new physics.
In this work, we simulate the light curve produced by a standard SN with a progenitor mass of $8.8~M_{\odot}$ and those produced by different BSM scenarios mentioned previously; see the left-hand panel of Fig.~\ref{fig:hits_and_likelihood} for light curves of the same two benchmarks from \cref{fig:fluxes}.

Since more than 93\% of the photons detected in IceCube originate in IBD interactions, the SM-only event rate peaks at around 0.1~s.
BSM scenarios, on the other hand, can give rise to an early or delayed signal depending on the particulars of the model.
In essence, Majorons can be generated before the peak of the SM $\bar{\nu}_{e}$ flux, and as light Majorons can travel nearly relativistically, they can escape the SNe with negligible time delay and then decay to an all-flavor flux of neutrinos, producing early $\bar{\nu}_{e}$ fluxes before the SM one.
The higher-mass Majorons, on the other hand, acquire a time delay from traveling sub-luminally and at a large detour, thus producing a later signal with sizable time delay.
We use the \texttt{ASTERIA}~\cite{spencer_griswold_2020_3926835} package to simulate the detector response to the different SN scenarios and the thermal and radioactive noise in the DOMs.
This package simulates light yields from coherent $\nu_{e}$ scattering off electrons in the ice, IBD of $\bar{\nu}_{e}$ on nuclei in the ice, and charged- and neutral-current interactions for $\nu_{\alpha}$. In addition to photons produced in neutrino interactions, it also simulates photons from thermal noise in the IceCube photomultipliers (PMTs) and from radioactive decay in the pressure glass housing.
In order to interface with \texttt{ASTERIA}, we use the \texttt{SNEWPY}~\cite{SNEWS:2021ewj} package and, in particular, the \texttt{ParametrizedFlux} object to parametrize the flux with the ``pinching'' factor~\cite{Keil:2002in,SNEWS:2021ewj}.

@@

\textbf{\textit{Results and Discussions}}---
\begin{figure}[t!]
    \centering
    \includegraphics[width=0.47\textwidth]{figures/majoran_sensitivity}
    \caption{\textbf{\textit{Exclusion sensitivities for the Majoron case.}}
    The lines on this plot show the exclusion sensitivity for IceCube, IceCube Gen-2, Hyper-Kamiokande and DUNE at $2\sigma$ CL.
    Additionally, we also show the reaches by SN cooling and by the lack of observation of high-energy neutrinos from SN1987A as black lines. 
    The sensitivities shown by the dashed line is obtained following~\cite{Brdar:2023tmi} with time-integrated data.
    }
    \label{fig:sensitivity}
\end{figure} 
Following our strategies above, we obtain the expected exclusion limits for various BSM scenarios from no observed excess of hits on top of background noise and standard neutrino flux.
The expected exclusion limits are presented in \cref{fig:sensitivity} for the Majoron case considering both IceCube and IceCube-Gen2. To illustrate the broad physics application of the analysis, we also show results for the sterile neutrino case with magnetic portal in the Supplemental Materials.
Current limits \cite{Fiorillo:2022cdq} from the energy loss requirement, and non-observation of high-energy SN neutrinos from SN 1987A are also shown in \cref{fig:sensitivity} for the Majoron case. With future SN explosions, IceCube can improve current limits by one order of magnitude. 
Moreover, while current limits from time-integrated observables follow the behavior of $g_\phi\propto m^{-1}_\phi$ as Majoron production rate is proportional to $(g_\phi m_\phi)^2$, IceCube can provide a stronger limit than this by focusing on the time window, especially before the peak of SM $\bar{\nu}_e$ fluxes, to reduce SM condemnations.

@@

There are existing uncertainties from different modeling of SN explosions (see e.g. Ref.~\cite{li2023old}) as well as discrepancies between SN 1987A neutrino data and modeling. Nevertheless, we point out that even the uncertainties of around $2\sigma$  will only change our exclusion limits by a small factor.
We make the code used in this work publicly available so that the impact of SN modeling uncertainties on the sensitivity of neutrino telescopes to BSM scenarios can be further explored.

\textbf{\textit{Summary}}---
In this Letter, we have demonstrated that neutrino telescopes such as IceCube can provide unprecedented probes of new physics, such as Majoron-like bosons and neutrino magnetic moment portal, by measuring SN time profiles to a window as short as $\mathcal{O}(0.01)$ s. Specifically, signatures outside the SM neutrino time window, including both early and delayed signals, can be effectively probed by neutrino telescopes. This analysis can be extended to test other BSM realizations including, but not limited to, sterile neutrinos with a mixing portal. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\textit{Acknowledgements}---
We would like to thank Thomas Janka and Daniel Kresse for providing the data from Garching Ssupernovae simulations (available at \url{https://wwwmpa.mpa-garching.mpg.de/ccsnarchive/}) in machine-readable form, and also for stimulating discussions. Y.-Y. L is supported by the NSF of China through Grant No. 12305107, 12247103.
CAA are supported by the Faculty of Arts and Sciences of Harvard University, the National Science Foundation (NSF), the NSF AI Institute for Artificial Intelligence and Fundamental Interactions (IAIFI), the Research Corporation for Science Advancement, 
and the David \& Lucile Packard Foundation.
CAA and JL were partially supported by the Alfred P. Sloan Foundation for part of this work.
JL acknowledges support from the Belgian American Educational Foundation.
